---
title: "APM1110_FA7_DeRamos_AneezaAnn_and_Eduvala_Mikhael"
author: "Aneeza Ann de Ramos and Mikhael Eduvala"
date: "2026-09-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

github link: https://github.com/AneezaAnn/APM1110_FA7_DeRamos_AneezaAnn_and_Eduvala_Mikhael.git

**The chosen scenario is how long one student tries entering the gate when their Face ID won't work at FEU gate 3. Here are the gathered data:**

1. 1 minute 23 seconds (83 seconds)

2. 0 minutes 54 seconds (54 seconds)

3. 0 minutes 51 seconds (51 seconds)

4. 1 minute 6 seconds (66 seconds)

5. 1 minute 21 seconds (81 seconds)

6. 0 minutes 42 seconds (42 seconds)

7. 0 minutes 24 seconds (24 seconds)

8. 0 minutes 38 seconds (38 seconds)

9. 0 minutes 26 seconds (26 seconds)

10. 0 minutes 31 seconds (31 seconds)

11. 0 minutes 52 seconds (52 seconds)

12. 0 minutes 17 seconds (17 seconds)

13. 0 minutes 48 seconds (48 seconds)

14. 0 minutes 22 seconds (22 seconds)

15. 0 minutes 10 seconds (10 seconds)

16. 1 minute 2 seconds (62 seconds)

17. 0 minutes 26 seconds (26 seconds)

18. 0 minutes 17 seconds (17 seconds)

19. 0 minutes 12 seconds (12 seconds)

20. 0 minutes 24 seconds (24 seconds)

21. 0 minutes 35 seconds (35 seconds)

22. 0 minutes 27 seconds (27 seconds)

23. 0 minutes 20 seconds (20 seconds)

24. 0 minutes 16 seconds (16 seconds)

25. 1 minute 14 seconds (74 seconds)

26. 0 minutes 33 seconds (33 seconds)

27. 0 minutes 21 seconds (21 seconds)

28. 0 minutes 25 seconds (25 seconds)

29. 0 minutes 18 seconds (18 seconds)

30. 0 minutes 14 seconds (14 seconds)

```{r}
# Vector of observed times in seconds

faceid_delay_sec <- c(83, 54, 51, 66, 81, 42, 24, 38, 26, 31,
                      52, 17, 48, 22, 10, 62, 26, 17, 12, 24,
                      35, 27, 20, 16, 74, 33, 21, 25, 18, 14)


# Calculate descriptive statistics

avg_delay_sec <- mean(faceid_delay_sec)
mins_component <- avg_delay_sec %/% 60
secs_component <- avg_delay_sec %% 60


# Model Parameter Calculation

rate_lambda <- 1 / avg_delay_sec


# Evaluate cumulative probability for t <= 30 seconds

target_cutoff <- 30
prob_within_cutoff <- pexp(target_cutoff, rate = rate_lambda)


# Output summary to console

cat(sprintf("Average Failure Duration: %d minute(s) and %.2f second(s)", mins_component, secs_component))
cat(sprintf("Rate Parameter (lambda): %.4f per second", rate_lambda))
cat(sprintf("Probability of resolving in <= 30 seconds: %.2f%%", prob_within_cutoff * 100))

```


**Analysis & Practical Applications**

While occasional biometric recognition issues seem minor in isolation, their cumulative impact during peak morning hours creates substantial pedestrian congestion. Our field observations indicate that when a scanner fails, a student spends up to 83 seconds attempting position adjustments or requesting manual clearance from security staff.

To mitigate these delays, administration should still implement a dual-authentication fallback—such (RFID tap card) and / or to take a new facial scan for those students who have trouble. These shall allow students to bypass stalled biometric turnstiles quickly.

**Summary of Results**

This investigation evaluated the operational impact of gate facial recognition errors across 30 sampled events. Standardizing all measurements into seconds yielded a sample mean ($\mu$) of 35.63 seconds per failed attempt. Fitting an exponential distribution model resulted in an estimated rate parameter $\lambda = 0.0281 \text{ s}^{-1}$. Using the cumulative distribution function (CDF), we determined that approximately 57.00% of failed entry attempts are cleared or abandoned within 30 seconds.


```{r}
# Re-assigning parameters for graphical plotting

rate_param <- 1 / mean(faceid_delay_sec)


# Constructing density histogram with theoretical overlay

hist(faceid_delay_sec,
     breaks = 9,
     freq = FALSE,
     col = "lightblue",
     border = "white",
     main = "Density Distribution of Face ID Entry Delays",
     xlab = "Duration of Failed Attempt (Seconds)",
     ylab = "Probability Density")


# Adding theoretical Exponential PDF Curve

curve(dexp(x, rate = rate_param),
      from = 0,
      to = max(faceid_delay_sec) + 10,
      col = "darkred",
      lwd = 2,
      add = TRUE)


# Legend positioning

legend("topright",
       legend = c("Observed Data", "Exponential Model"),
       fill = c("lightblue", NA),
       col = c(NA, "darkred"),
       lwd = c(NA, 2),
       bty = "n")
```
